The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-05-03 Paper.Rmd
making another plot for main text

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---
updated on `r Sys.Date()`

# Front matter

@@

Left- and right-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
are related in the left and right tails, respectively, of their 
distributions [@nelsen2007_copula]. 
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions.

![Schematic diagram supporting defintions of nonparametric tail dependence statistics.\label{fig_stat_image}](./stat_image.jpg){ width=25%, height=25% }

@@

In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
$\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated
from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).

>>>>

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---
updated on `r Sys.Date()`
```{r setup_Paper, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
seed<-101
# families<-c(1,3:10,13,14,16:20)
source("mtime.R") #A function needed for caching
```
# Front matter

@@

Left- and right-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
are related in the left and right tails, respectively, of their 
distributions [@nelsen2007_copula]. 
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Fig. \ref{fig_stat_image}(A)). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions.

```{r fig_stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
set.seed(seed) 
source("fig_Paper_stat_testing.R")
load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
resultsloc<-"./Results/stat_results/stat_testing/"

copname<-"C"
reslist<-res_pt35_rankT

statname<-"CorlmCoru"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
statname<-"PlmPu"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
statname<-"D2umD2l"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
```

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{5.6 cm} (B)} \\
\vspace{-0.6 cm}
\begin{center}
\hspace{0.5 cm}
\includegraphics[width=5 cm,height=4.3 cm]{./stat_image.jpg} \hspace{0.6 cm}
\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/CorlmCoru_C.pdf}
\end{center}
\textbf{ \hspace{6 cm} (C) \hspace{5.7 cm} (D)} \\
\vspace{-0.6 cm}
\begin{center}
\hspace{0.2 cm}
\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/PlmPu_C.pdf} \hspace{0.3 cm}
\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/D2umD2l_C.pdf}
\caption{(A) Schematic diagram supporting defintions of nonparametric tail dependence statistics; (B) Results of tests of $\cor_l-\cor_u$, (C) $\Ps_l-\Ps_u$, and (D) $\Dsq_u-\Dsq_l$ reveal the known asymmetry (more lower tail than upper) of tail dependence of the Clayton copula with $n=35$ data points. \label{fig_stat_image}}
\end{center}
\end{figure}

@@

In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
also tested our statistics in Fig. \ref{fig_stat_image}(B - D) by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
$\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated
from Clayton copulas (for details see Appendix \ref{SM-Test_npa_stat}).
